---
title: "README.Rmd"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Overview

"Binomial" is a package that implements functions for calculating probabilities of a Binomial random variable, and related calculations such as the probability distribution, the expected value, variance, etc.
Some examples of these functions are:

  * 'bin_distribution()' creates a dataframe (of class 'binomial distribution')
  * 'bin_cumulative()' creates a dataframe (of class binomial cumulative distribution)
  * 'plot()' method for a 'bin_distribution' and 'bin_cumulative' object to plot probabilities of successes.
  * 'bin_variable()' creates a list object (of class binomial random variable)
  * 'summary()' method for a binomial variable object
  
## Motivation

This package was created as a part of Workout3 for Stat 133.

## Installation

Install the development version from GitHub via the package `"devtools"`:

```{r}
# development version from GitHub:
#install.packages("devtools") 
# install "cointoss" (without vignettes)
devtools::install_github("divyavijay14/stat133-spring19/Binomial")
# install "cointoss" (with vignettes)
devtools::install_github("divyavijay14/stat133-spring19/Binomial", build_vignettes = TRUE)
```

## Usage

```{r}
library(Binomial)

bin_probability(success = 2, trials = 5, prob = 0.5)

#creates a dataframe of probabilities of successes
bin_distribution(trials = 5, prob = 0.5)

#plots that binomial distribution
dis1 <- bin_distribution(trials = 5, prob = 0.5)
plot(dis1)

#creates a dataframe of cumulative probabilities of successes
bin_cumulative(trials = 5, prob = 0.5)

#plots that cumulative binomial distribution
dis2 <- bin_cumulative(trials = 5, prob = 0.5)
plot(dis2)

#gives us summary statistics on this binomial random variable
bin1 <- bin_variable(trials = 5, p = 0.5)
binsum1 <- summary(bin1)
binsum1

```

